Using Genetic Algorithms to Find Approximations for the Minimum Vertex Cover Problem

### 2018-03-19 paper.tex
added amareen notes

They considered regular graphs, each of which consists of three levels.
The first two levels have the same
number of nodes while the third level has two nodes less
than the number of nodes found on the previous two
levels.
More precisely, each graph consists of $n = 3k + 4 \ (k \geq 1)$
nodes; $k+2$ nodes on the first level, labeled $1, 2, \dots, k+2$;
followed by $k+2$ nodes on the second level, labeled $k+3, \dots, 2k+4$;
while the $k$ nodes of the third level are labeled $2k+5, \dots, 3k+4$.
The regular graph for $k = 3$ can be found in figure~\ref{fig:ps}.
A minimum vertex cover is obtained by choosing all
the nodes of the second level, but the greedy strategy
would start with the nodes of the third level, since these
have highest degree. Consequently, the greedy strategy
finds a solution of size $2k + 2$ (since it has to select $k + 2$
nodes in addition to the nodes of the third level), while
the optimal solution has size $k + 2$~\cite{mvcp-back}.

However this algorithm is not totally useless. It may be shown that it always
achieves the ratio $\Theta(\lg n)$\footnote[2]{$f(n) = \Theta(g(n)) \iff
\exists \alpha, \beta, n_0 : \forall n \geq n_0, \ 
\alpha \cdot g(n) \leq f(n) \leq \beta \cdot g(n)$ \cite[pp.~41--43]{cormen}},
where $n$ is the number of vertices.
That is, if $C^*(n)$ is the size of the optimal
solution and $C(n)$ is the size of the solution found by the greedy
algorithm. Then we have $\frac{C(n)}{C^*(n)} = \Theta(\lg n)$ \cite[p.~323]{9:mvcp}.

@@

The term $ \sum_{i=1}^{n} x_i $ of $ f(\vec{x}) $ determines the size of the
potential vertex cover represented by $\vec{x}$, while the term
$ n \cdot \sum_{i=1}^{n} \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij} $
penalizes sets $V'$ that
are not covers by adding a penalty of magnitude $n$ for
each edge $e_{ij}$ for which $i \notin V' \wedge j \notin V'$
(note that $e_{ii}$ is previously defined to be zero).
Consequently, any infeasible solution will have a fitness greater than
or equal to $n$. And for any feasible solution the second term will
drop to zero, making the fitness value of the worst feasible solution at most $n$.

@@

\item The penalty should be graded, i.e., fitness values
should improve as solutions approach feasible regions
of the search space. In other words, the more edges a solution
covers, the fitter it is. notice that this is guaranteed by the
second term in equation~\ref{eq:f}, where for each uncovered edge
the fitness is penalized with the value of $n$.
\end{itemize}

@@

\paragraph{Test set~2:}
To test the behaviour of the genetic algorithm
as well as the vercov heuristic for an even
larger problem size, The same experiments
were performed for larger graphs.
The results in table~\ref{tbl:t3}
and table~\ref{tbl:t4} correspond to test set~2,
which consists of five random graphs of size
$n = 200$ with different edge densities $d$.
Densities $d \in \{0.1, 0.2, 0.3, 0.4, 0.5\}$ are used.
Moreover, we construct the graphs in a way that guarantees an optimum of quality 110
(see algorithm~\ref{alg:gen}, we set $k = 110$).
In addition to a graph instance of the regular graphs introduced by
Papadimitriou and Steiglitzof of size $n = 220 \ (k=66)$.
In this case, the genetic algorithm
was allowed to run for $4 \cdot 10^4$ function evaluations.\\
The corresponding results obtained by the
genetic algorithm are shown in table~\ref{tbl:t3}, while the results
from the vercov heuristic are shown in table~\ref{tbl:t4}.
See the previous paragraphs for a better understanding of these results.

@@

Moreover, the randomly constructed problems quickly
become simpler for the genetic algorithm when the edge
density is increased.
For an edge density of $d \in \{0.4, 0.5\}$, the
genetic algorithm almost surely finds the global optimum
of the problems, while vercov heuristic gives
a relatively terrible results.

@@

For the randomly constructed graphs, the vercov heuristic
performs poorly.
It barely finds solutions of quality 80 (for graphs of size 100)
and 172 (for graphs of size 200).
Moreover, notice that the quality of its solutions are dominated
by the quality of the solutions found by the genetic algorithm;
the worst solutions found by the genetic algorithm are almost as
good as the best solutions found by the vercov heuristic.

@@

\section{Conclusion}
\label{sec:conc}
This work has demonstrated that genetic algorithms can
be used in a fairly straightforward way to find good
approximative solutions for NP-hard problems, namely
the minimum vertex cover problem.
The robustness of the presented solution along with
the positive results obtained support the idea
that genetic algorithms are a desirable approach for tackling
these type of problems.
To tackle a new problem, rather than having to construct
tailored heuristics to handle the problem under consideration,
we advocate the use of genetic algorithms where the
only change to perform is the formulation of a new fitness
function.

>>>>

They considered regular graphs, each of which consists of three levels.
The first two levels have the same
number of nodes while the third level has two nodes less
than the number of nodes found on the previous two
levels.
More precisely, each graph consists of $n = 3k + 4 \ (k \geq 1)$
nodes; $k+2$ nodes on the first level, labeled $1, 2, \dots, k+2$;
followed by $k+2$ nodes on the second level, labeled $k+3, \dots, 2k+4$;
while the $k$ nodes of the third level are labeled $2k+5, \dots, 3k+4$.
The regular graph for $k = 3$ can be found in figure~\ref{fig:ps}.
A minimum vertex cover is obtained by choosing all
the nodes of the second level, but the greedy strategy
would start with the nodes of the third level, since these
have the highest degree. Consequently, the greedy strategy
finds a solution of size $2k + 2$ (since it has to select $k + 2$
nodes in addition to the nodes of the third level), while
the optimal solution has size $k + 2$~\cite{mvcp-back}.

However, this algorithm is not totally useless. It may be shown that it always
achieves the ratio $\Theta(\lg n)$\footnote[2]{$f(n) = \Theta(g(n)) \iff
\exists \alpha, \beta, n_0 : \forall n \geq n_0, \ 
\alpha \cdot g(n) \leq f(n) \leq \beta \cdot g(n)$ \cite[pp.~41--43]{cormen}},
where $n$ is the number of vertices.
That is, if $C^*(n)$ is the size of the optimal
solution and $C(n)$ is the size of the solution found by the greedy
algorithm. Then we have $\frac{C(n)}{C^*(n)} = \Theta(\lg n)$ \cite[p.~323]{9:mvcp}.

@@

The term $ \sum_{i=1}^{n} x_i $ of $ f(\vec{x}) $ determines the size of the
potential vertex cover represented by $\vec{x}$, while the term
$ n \cdot \sum_{i=1}^{n} \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij} $
penalizes sets $V'$ that
are not covers by adding a penalty of magnitude $n$ for
each edge $e_{ij}$ for which $i \notin V' \wedge j \notin V'$
(note that $e_{ii}$ is previously defined to be zero).
Consequently, any infeasible solution will have a fitness greater than
or equal to $n$. And for any feasible solution, the second term will
drop to zero, making the fitness value of the worst feasible solution at most $n$.

@@

\item The penalty should be graded, i.e., fitness values
should improve as solutions approach feasible regions
of the search space. In other words, the more edges a solution
covers, the fitter it is. Notice that this is guaranteed by the
second term in equation~\ref{eq:f}, where for each uncovered edge
the fitness is penalized with the value of $n$.
\end{itemize}

@@

\paragraph{Test set~2:}
To test the behaviour of the genetic algorithm
as well as the vercov heuristic for an even
larger problem size, the same experiments
were performed for larger graphs.
The results in table~\ref{tbl:t3}
and table~\ref{tbl:t4} correspond to test set~2,
which consists of five random graphs of size
$n = 200$ with different edge densities $d$.
Densities $d \in \{0.1, 0.2, 0.3, 0.4, 0.5\}$ are used.
Moreover, we construct the graphs in a way that guarantees an optimum of quality 110
(see algorithm~\ref{alg:gen}, we set $k = 110$).
In addition to a graph instance of the regular graphs introduced by
Papadimitriou and Steiglitzof of size $n = 220 \ (k=66)$.
In this case, the genetic algorithm
was allowed to run for $4 \cdot 10^4$ function evaluations.\\
The corresponding results obtained by the
genetic algorithm are shown in table~\ref{tbl:t3}, while the results
from the vercov heuristic are shown in table~\ref{tbl:t4}.
See the previous paragraphs for a better understanding of these results.

@@

Moreover, the randomly constructed problems quickly
become simpler for the genetic algorithm when the edge
density is increased.
For an edge density of $d \in \{0.4, 0.5\}$, the
genetic algorithm almost surely finds the global optimum
of the problems, while vercov heuristic gives relatively terrible results.

@@

For the randomly constructed graphs, the vercov heuristic
performs poorly.
It barely finds solutions of quality 80 (for graphs of size 100)
and 172 (for graphs of size 200).
Moreover, notice that the quality of its solutions is dominated
by the quality of the solutions found by the genetic algorithm;
the worst solutions found by the genetic algorithm are almost as
good as the best solutions found by the vercov heuristic.

@@

\section{Conclusion}
\label{sec:conc}
This work has demonstrated that genetic algorithms can
be used in a fairly straightforward way to find good
approximative solutions for NP-hard problems, namely
the minimum vertex cover problem.
The robustness of the presented solution along with
the positive results obtained support the idea
that genetic algorithms are a desirable approach for tackling
these types of problems.
To tackle a new problem, rather than having to construct
tailored heuristics to handle the problem under consideration,
we advocate the use of genetic algorithms where the
only change to perform is the formulation of a new fitness
function.
